---
title: "p8105_hw3_yz4433"
author: "Yifei Zhao"
date: "2022-10-13"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)
library(stringr)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1
### description of data
```{r}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)
```
This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```
In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.

### plot of items
```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

## Problem 2
### Manipulate data
```{r}
data_acc = read_csv(file = "./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  add_column(group = "days", .after = "day") %>% 
  mutate(group = ifelse((day == "Saturday" | day == "Sunday"), 'weekend', 'weekday'))
```
The final `r nrow(data_acc)` $\times$ `r ncol(data_acc)` dataset, data_acc,  illustrates activity counts for each minute of a 24-hour day starting at midnight of 35 days in 5 weeks. Parts of counts values are 1 and the dataset includes a weekday vs weekend variable.

### Table showing totals
```{r}
data_acc = data_acc %>%
  add_column(total = rowSums(data_acc[,5:1444]), .after = "group")
knitr::kable(data_acc[, 1:5], "simple")
```
According to the table, we find the maximum is `r max(data_acc$total)` and the minimum is `r min(data_acc$total)`, but no apparent trends relating to weekdays or weekends (certain days in general week).

### Activity plot
```{r}
data_acc = data_acc[,-5] %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity", 
    values_to = "act_counts") %>%
  mutate(activity = as.integer(substr(activity, 10, 13)))
```

```{r}
ggplot(data_acc, aes(x = activity, y = act_counts)) + 
  geom_point(aes(color = day), size = 1, alpha = 0.5) +
  ggtitle("24-hour activity time courses for each day") +
  xlab('activity time in minute') +
  scale_x_continuous(breaks = seq(0, 1440, 60)) +
  theme(axis.text.x = element_text(angle = 90))
```
From the activity plot, the data have several peaks and troughs. Overall, Monday high Wednesday low. There are two main peak areas, activity time points lie in (570,750) and (1170,1320), and the main trough area is about (60,240). The results are strongly relating to people's daily life (low activity when sleep, high when work or study).

## Problem 3
### Data cleaning
```{r}
data("ny_noaa")
ny_noaa = ny_noaa %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("year", "month", "day")) %>% 
  mutate(prcp = as.numeric(prcp)/10, tmax = as.numeric(tmax)/10, tmin = as.numeric(tmin)/10, snow = as.numeric(snow))
```

```{r}
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}
```
We divide the prcp, tmax, and tmin values by 10 to get the real precipitation and temperature values. The mode of snowfall value is `r getmode(ny_noaa$snow)`. The reason could be that temperatures and precipitation of the data imply not strong conditions of snowfall in most cases.

### temperature plot
#### Jan and July data
```{r}
janmax =
  ny_noaa %>%
  drop_na(tmax) %>%
  filter(month == "01")
df_jan = aggregate(tmax ~ id+year, janmax, mean)

julmax =
  ny_noaa %>%
  drop_na(tmax) %>%
  filter(month == "07")
df_jul = aggregate(tmax ~ id+year, julmax, mean)
```

#### plotting
```{r}
tmaxjan = 
  df_jan %>% 
  ggplot(aes(x = year, y = tmax, color = id, group = id)) + 
  ggtitle(" Average max temperature in January") +
  geom_point(size = 1, alpha = .5) +
  geom_line() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(legend.position = "none")
```

```{r}
tmaxjul = 
  df_jul %>% 
  ggplot(aes(x = year, y = tmax, color = id,  group = id)) + 
  ggtitle(" Average max temperature in July") +
  geom_point(size = 1, alpha = .5) +
  geom_line() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(legend.position = "none")

tmaxjan/tmaxjul
```
According to the year-tmax 2-panel plot, the max temperatures in January fluctuate greatly, and these two subplots are of the similar trends (correspondingly reach the peaks and troughs), which implies the seasonal or yearly weather simultaneously influence the near dates. Both of the data have outliers. For example, the lowest point in Jan 1982 and the lowest point in Jul 1988.

### tmax vs tmin, and snowfall plotting
```{r}
ny_temp = 
  ny_noaa %>% 
  drop_na(tmax) %>% 
  drop_na(tmin) %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  ggtitle("tmax vs tmin") +
  geom_hex(alpha = .5) +
  theme(legend.position = "bottom", legend.key.size = unit(30, "pt"))
  # theme(axis.text.x=element_blank())
  # theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r}
ny_snow =
  ny_noaa %>%
  filter(snow > 0 & snow < 100) %>% 
  ggplot(aes(x = snow, color = year)) +
  ggtitle("distribution of snowfall values") +
  geom_density(alpha = .5) + 
  viridis::scale_fill_viridis(discrete = TRUE) +
  theme(legend.position = "bottom", legend.key.size = unit(6, "pt"))
  # theme(axis.text.x=element_blank())
  # theme(axis.text.x = element_text(angle = 90, hjust = 1))

ny_temp + ny_snow
```